---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# solong

[![Travis-CI Build Status](https://travis-ci.org/SCAR/solong.svg?branch=master)](https://travis-ci.org/SCAR/solong)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/SCAR/solong?branch=master&svg=true)](https://ci.appveyor.com/project/SCAR/solong)
[![codecov](https://codecov.io/gh/SCAR/solong/branch/master/graph/badge.svg)](https://codecov.io/gh/SCAR/solong)

Overview
--------

This R package provides allometric equations that relate the body size of Southern Ocean taxa to their body part measurements.

## Packaged equations

```{r include=FALSE}
devtools::load_all()
library(dplyr)
```
The package currently includes `r nrow(sol_equations())` equations, covering mostly cephalopods and fish. A breakdown of the number of equations by taxonomic class and the allometric property that they estimate:

```{r echo=FALSE,cache=TRUE}
library(worrms)
uid <- unique(sol_equations()$taxon_aphia_id)
ucl <- lapply(uid,wm_classification)
uid_cls <- tibble(taxon_aphia_id=uid,taxonomic_class=vapply(ucl,function(z)z$scientificname[z$rank=="Class"],FUN.VALUE=""))
```
```{r echo=FALSE}
knitr::kable(sol_equations() %>% left_join(uid_cls,by="taxon_aphia_id") %>% xtabs(~taxonomic_class+return_property,data=.))
```


Installing
----------

```{r, eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("SCAR/solong")
```

Usage
-----

```{r eval=FALSE,message=FALSE, warning=FALSE}
library(solong)
library(dplyr)
```


Let's say we have some measurements of *Architeuthis dux* squid beaks:

```{r}
x <- tibble(LRL=c(11.3,13.9),species=c("Architeuthis dux"))
x
```

It doesn't matter what the column names are, but we do need to set the properties of the columns so that `solong` can find the appropriate data to use in each allometric equation. Here we've measured lower rostral length, so:

```{r}
x$LRL <- sol_set_property(x$LRL,"lower rostral length")
```

Now we can apply allometric equations to our data. What equations do we have available for our species of interest?

```{r}
subset(sol_equations(),taxon_name=="Architeuthis dux") %>% summary
```

Here we use the equation with ID `342218_ML_Roel2000`, which is from Roeleveld (2000) and gives the mantle length of *Architeuthis dux* based on the lower rostral length.

This equation can be applied to to all rows:
```{r}
sol_allometry(x,c("342218_ML_Roel2000"))
```

Or we can apply a different equation to each row. Here we could use different allometric equations for mantle length:
```{r}
xa <- sol_allometry(x,c("342218_ML_Roel2000","342218_ML_Clar1986"))
xa
```

The `allometric_value` column contains the values that have been estimated, and the `allometric_property` column gives the name of the property that has been estimated.

We can also see that the returned `allometric_value` is of that specific property, with appropriate units:
```{r}
sol_get_property(xa$allometric_value)

units(xa$allometric_value)
```

## Details

We can apply equations that use different inputs, provided that they estimate the same output property. For example, equation `342218_WW_Clar1986` estimates the body weight of the squid *Architeuthis dux* based on lower rostral length measurements:

```{r}
sol_equation("342218_WW_Clar1986") %>% summary
```

And equation `195932_WW_GaBu1988` estimates the weight of male Weddell seals based on their standard length:

```{r}
sol_equation("195932_WW_GaBu1988") %>% summary
```
Note that this equation estimates weight in kg, whereas `342218_WW_Clar1986` estimates weight in g. We can apply the two equations together to a single data set:

```{r}
x <- tibble(LRL=c(11.3,NA),species=c("Architeuthis dux","Leptonychotes weddellii"),SL=c(NA,175)) %>%
  mutate(LRL=sol_set_property(LRL,"lower rostral length"),
         SL=sol_set_property(SL,"standard length","cm"))

xa <- sol_allometry(x,c("342218_WW_Clar1986","195932_WW_GaBu1988"))

xa %>% select(species,allometric_property,allometric_value)
```

The output values are of property "wet weight" and have all been provided in g (because the output column `allometric value` must have a single set of units):

```{r}
sol_get_property(xa$allometric_value)

units(xa$allometric_value)
```


If we try to apply equations that estimate different properties, we will get a warning:
```{r}
x <- tibble(LRL=c(11.3,13.9),species=c("Architeuthis dux")) %>%
  mutate(LRL=sol_set_property(LRL,"lower rostral length"))

xa <- sol_allometry(x,c("342218_ML_Roel2000","342218_WW_Clar1986"))

xa %>% select(species,allometric_property,allometric_value)
```

And while the `allometric_property` column still says which property was estimated for each row, the property type and units of the returned `allometric_value` will not be set, because they are not consistent across the different equations:
```{r}
sol_get_property(xa$allometric_value)
```

### Missing information

What happens if we don't have the required information in our data to use a particular equation? The `234631_SL~OL_WiMc1990` equation is for fish length, and requires otolith length (not present in our test data).

```{r}
tryCatch(
  sol_allometry(x,"234631_SL~OL_WiMc1990"),
  error=function(e) conditionMessage(e)
)
```

### Reliability of equations

Older equations were typically published along with the number of samples used to fit the equation (N) and the resulting goodness-of-fit of the equation to the data (R^2). These two quantities (if provided by the original source) can be found in the `reliability` component of an equation, and can be used to help decide if a given equation is appropriate for your data.

Some equations, typically from more recent publications, also provide the standard errors of the coefficients (or similar information) and thereby allow the `allometric_value_lower` and `allometric_value_upper` values to be estimated (the upper and lower bounds on the estimate). These should give a more reliable indicator of the precision of the estimated quantities.

```{r}
x <- tibble(TL=c(10) %>% sol_set_property("carapace length",with_units="mm"))
sol_allometry(x,"369214_WW_Lake2003") %>%
  select(allometric_value,allometric_value_lower,allometric_value_upper)
```

Attempts are made to avoid allowing an equation to extrapolate beyond its valid input data range. Some equations will explicitly return `NA` results for such inputs. The `inputs` component of the equation may also hold information about the range of the inputs used to fit the equation, which may help assess whether your data lie within its valid range.

### Other sources of equations

TODO: document the `sol_fb_length_weight()` function for using FishBase length-weight equation coefficients.

### Adding your own equations

TODO: document, including what to do when a property is not part of the `sol_properties()` collection.

### Taxonomy

Equations are registered against *taxon_name* and *taxon_aphia_id* (the species identifier in the World Register of Marine Species). The *taxon_aphia_id* may be more reliable than species names, which can change over time. Users might like to look at the [worrms package](https://cran.r-project.org/package=worrms) for interacting with the World Register of Marine Species.


## Related packages


- [units](https://cran.r-project.org/package=units) for handling units of measurement
- [worrms](https://cran.r-project.org/package=worrms) for taxonomy
- [shapeR](https://cran.r-project.org/package=shapeR) for collection and analysis of otolith shape data
